%!TEX root = ../../tese.tex
%!TEX encoding = UTF-8 Unicode
\raggedbottom
\pdfbookmark{Acknowledgements}{Acknowledgements}
\begin{acknowledgments}
\begin{doublespace}

This dissertation owes its existence from the proposal made by my advisor, Miguel Pupo Correia, who supports me since the beginning, allowed me to do this thesis by distance and gave me many great ideas and enthusiasm.\\
Thanks to the creators of my Family (my grand-parents), specially to my grand-father José Paulo Nascimento who became an orphan when he was young and humbly worked hard on civil construction, agriculture and aviculture to create a big happy and reliable family. Thanks for your great inspiration.
Equal thanks go to my parents for supporting all my crazy life decisions, great education and happy family environment. Thanks to my sister and cousins for so much fun and good moments.
Francisco Malheiro, Mario Malheiro, Teresa Alves, Sara Alves, Paula Alves, Pedro Alves, André Santos, Joana Malheiro my hearth friends since ever.\\
Pedro Gomes, Marcella Monteiro (special gratefulness for many English texts reviews), Vanessa Silva, Rui Caldas, Tânia Rodrigues my crazy and great friends since high-school. Gonçalo Pestana (great mate in Aalto), Artur Balanuta, David Dias, Gonçalo Carito, Francisco Esteves, Tiago Freire, Rita Raimundo, Bruno Mercês, Paulo Alves, Catarina Moura my closest friends in Instituto Superior Técnico.\\
Thanks go as well to all my friends from BEST, Erasmus, Filarmónica da Encarnação, Banda Fórum and the startups where I worked: NWC and Eduze who tought me what universities can’t. All my BSc. and MSc. teachers for high competency and incredible skills. Special acknowledgement to Prof. Rui Valadas, my wise studies advisor who gave me a lot of good advice.\\
Special thanks to Maike Marzendorfer for being always by my side sharing best and worst moments of my life.\\
This work was supported by INESC-ID - Distributed Systems group and part of RC-Clouds - Resilient Computing in Clouds - research project funded by FCT- Fundação para a Ciência e a Tecnologia (PTDC/EIA-EIA/115211/2009). 
%The information presented in this dissertation does not necessarily reflect the position or the policy of these sponsors and no official endorsement should be inferred.

\end{doublespace}

\begin{flushright}
   
\begin{tabular}{r}
 \\ Lisbon, \monthYearDate\today \\
Dário Nascimento\\
\end{tabular}
\end{flushright}


\end{acknowledgments}
\clearpage
\thispagestyle{empty}
\cleardoublepage